Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2020-05-07 manuscript/manuscript.tex
more Gs changes

@@

>>>>

\subsubsection{Truncation Selection.}
\gc{We explored how directional truncation selection generated
  temporal covariances.In these simulations we select the top ${\bf X
    \%}$ of the phenotypic distribution of individuals to form the
  next generation. We burnin these simulations {\bf by
    XXXX}. More extreme directional truncation selection generated
  larger initial covariances (Figure
  \ref{suppfig:supp-trunc}B). However, weaker truncation selection
  generated more sustained positive covariances and so had a larger
  long term impact

@@

% \subsubsection{The Effect of Fixations}
% \label{supp:fixation}

% In analyzing both the empirical datasets included in our study and our
% simulation results, we noticed the temporal covariances and $G(t)$ statistics
% can differ depending on how allele frequencies of zero or one are handled.
% Generally, temporal covariances should be calculated on polymorphic sites; once
% a site has reached fixation or loss, its allele frequency change $\Delta p_t =
% 0$ and including these sites in the temporal covariance calculation can lead to
% biases, which we discuss below. However, with \emph{sample} allele frequencies,
% rather than population frequencies, a site with observed frequency zero or one
% may still be segregating, but by chance not sampled at a timepoint. Here, we
% discuss the effect of including sites with frequency zero or one, and show our
% empirical results are not qualitatively different when analyzed excluding fixed
% sites.

% \begin{figure}[!ht]
%   \centering
%   \includegraphics[width=\textwidth]{figures/fig-bgs-G-with-fixations.pdf}

%   \caption{$G(t)$ trajectories from background selection simulations both
%   including fixed sites (solid lines) and not including fixed sites (dashed
% lines). The dashed lines are identical to those in Supplementary Materials
% Figure \ref{suppfig:supp-bgs-g}. }

%   \label{suppfig:supp-bgs-g-fix}
% \end{figure}

% We noticed in some simulations, when temporal covariances are calculated over
% longer time periods (i.e. $>50$ generations), that neutral simulations appear
% to slowly trend towards negative $G(t)$ (Supplementary Materials Figure
% \ref{suppfig:supp-bgs-g-fix}). In investigating this bias, we found that the
% effect went away when we marked fixed or lost sites as missing, and calculated
% the temporal covariance; furthermore, this affected the $G(t)$ trajectories
% under selection too (Supplementary Material Figure
% \ref{suppfig:supp-bgs-g-fix}).

% Looking a neutral simulations results, we found that spuriously significant
% $G(t)$ trajectories (e.g. different from zero under the case of neutrality) can
% be created by conditioning on sojourn time. Our intuition, which will need to
% be confirmed by future work and theory, is that by conditioning on a quick
% sojourn time to fixation (or loss) from an initial intermediate frequency,
% leads to trajectories that, purely by chance, have many upward (or downward)
% allele frequency changes, which leads to a positive sum of pairwise temporal
% covariances (the numerator of $G(t)$). Similarly, conditioning on long sojourn
% times implies that frequency changes alternate back and forth quite a bit, such
% the sum of pairwise temporal covariances could be slightly negative. However,
% the overall affect could also be affected by the average allele frequency
% change across loci and other factors.

% To further complicate the matter, our simulation allele frequencies are
% population, rather than sample frequencies, when an allele is fixed or lost is
% known with certainty. With empirical data calculated on sample allele
% frequencies, low frequency minor alleles many not be sampled at some
% timepoints, and excluding these observations (instead of treating it as a
% trajectory that has a 0 frequency timepoint) biases estimates.  We observed
% this bias created by ignoring fixed or loss sites by comparing our $N_e$
% estimates for the \textcite{Barghi2019-qy} study (estimated from the total
% variance in allele frequency change) with the estimates from the original
% paper. We found that ignoring fixed or lost sites (by treating them as missing
% data in calculating the covariances) led to many low-frequency alleles not
% contributing to the variance, leading this statistic to be calculated on more
% intermediate frequency alleles and thus biasing the $N_e$ estimate downwards.
% Additionally, we tried only dropping fixed or loss sites from the temporal
% covariance calculations that were at the end or the beginning of a trajectory
% (e.g., as if the site was created by a new mutation or fixed); while this
% ameliorated some of this bias, it still did not lead to $N_e$ estimates
% congruent with the original studies. Overall, we found by trying all these
% approaches that not removing fixed or lost sites was the best way to deal with
% sample allele frequencies that could be missing from some timepoints.

% \begin{figure}[!ht]
%   \centering
%   \includegraphics[width=0.6\textwidth]{figures/figure-1-G-covs-nofix.pdf}

%   \caption{The effect of excluding fixed/lost sites in the calculation of the
%   temporal covariances and $G(t)$ trajectories of the \textcite{Barghi2019-qy}
% data. Dashed lines are those including fixed/lost sites (i.e. the original
% Figure 1), and solid lines are excluding fixed/lost sites.}

%   \label{suppfig:supp-fig-1-nofix}
% \end{figure}

% \begin{figure}[!ht]
%   \centering
%   \includegraphics{figures/figure-3-hists-b-without-fixations.pdf}

%   \caption{A version of Figure \ref{fig:figure-3} (A) and (B) excluding fixed
%     and lost sites. The same qualitative pattern holds as the original figure,
%     which did not exclude fixed and lost sites: there is an enrichment of
%     positive temporal covariances between near timepoints ($k=2$) in the
%   \textcite{Barghi2019-qy} study, and an excess of negative temporal
% covariances at more distant timepoints ($k=2$). }

%   \label{suppfig:supp-fig-3-nofix}
% \end{figure}

% To ensure that our findings were robust to handling sites with a frequency of
% zero or one differently, we regenerated Figures \ref{fig:figure-1} and Figure
% \ref{fig:figure-3} but excluded frequencies of zero or one. Specifically, we
% wanted to ensure that our finding that later timepoints had negative
% covariances at later timepoints was not spuriously caused by the way fixed or
% lost sites are handled. We see no qualitative difference (Supplementary
% Material Figures \ref{suppfig:supp-fig-1-nofix} and
% \ref{suppfig:supp-fig-3-nofix}) that emerges when sites with frequency zero or
% one are excluded.

% \begin{figure}[!ht]
%   \centering
%   \includegraphics[width=\textwidth]{figures/fig-both-trunc.pdf}

%   \caption{$G(t)$ trajectories (A) and temporal covariances (B) from
%     truncation selection simulations for different numbers of individuals
%     selected (line color). Dashed lines indicate $G(t)$ trajectories and
%     temporal covariances calculated \emph{including} fixed sites, while the
%     solid lines exclude fixed sites. All values are averaged over 30 replicate
%     simulations; the lines in the right figure are loess smoothed, while points
%     are averages. The solid lines of the temporal covariances have been
%     excluded in the left figure for clarity, but are similar except they do not
%     become negative.}

%     \label{suppfig:supp-trunc}
% \end{figure}

% Through simulations, we noticed the strongest effect fixation  or loss of sites
% has on temporal covariance and $G(t)$ is under truncation selection. Here,
% since only a (potentially small) fraction of individuals contribute to the next
% generation, sites can fix over very short timescales. Furthermore, the small
% number of effective breeders contributing to the next generation shrinks $N_e$
% considerably, which increases $\var(p_t - p_0$, the denominator of $G(t)$. We
% see the effect of handling fixed/lost sites differently, and the faster rate of
% drift in Supplementary Materials Figure \ref{suppfig:supp-trunc} (A), where
% weaker truncation selection actually has higher levels of $G(t)$. Looking just
% at temporal covariances, we find that stronger truncation selection (e.g.  a
% smaller tail of individuals selected) does lead to greater temporal
% covariances.
